The pragmatics of perspective-taking

### 2018-03-20 writing/journal_manuscript/manuscript.tex
Streamline Exp. 1; start Exp. 2

%While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \cite{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}. 
We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \cite{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context. 
For example, suppose a friend wants to refer to a nearby Dalmatian. 
%There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \cite{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}. 
When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.

@@

@@

%To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start. 
Participants were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig. \ref{fig:interface}). Following \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, five of the grid cells were occluded from the director's perspective, and six or seven objects were displayed in the grid at a given time. On each trial, the director was shown an instruction indicating how one of the objects should be moved. They then used the chatbox to send a message to their partner  The experiment was composed of eight items, with each item using a different set of objects. Each item included one `critical pair' of objects, one of them the target and the other hidden, such as the cassette tape and the roll of tape. For each item, we gave the director a series of four instructions to move objects around, which were displayed as a series of arrows pointing from some object to an unoccupied cell.

Participants could freely type messages to one another at any point.

One of these objects was a `target', such as a cassette tape, placed in an unoccluded cell such that both participants could see it. Another object, such as a roll of tape, was placed in an occluded slot such that it was only visible to the matcher. The rest of the objects were unrelated `fillers' placed in random locations. We used the same set of targets and occluded alternatives as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, but we were unable to obtain the filler objects from the original experiment and created our own.

Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface. %Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.

To collect clean mouse-tracking data, we began every instruction by asking the matcher to click a small circle in the center of the grid. After this small circle was clicked, the director was allowed to communicate the next instruction and we started recording from the matcher's mouse. 
One of the instructions was a `critical instruction,' which referred to the target object.
For half the instructions, directors were free to communicate however they wished. For the other half, including all the critical instructions, their messages to the matcher were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat window, and they would have to click `send' for the experiment to continue.)

We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Each participant received half the items in the experimental condition and half in the baseline condition. The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items. All object sets, object placements, and corresponding instruction sets were the same for all participants.

This paradigm differs from those used by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} in three primary ways. First, participants were not seated across from each other at a table: they each saw a view of the 4 x 4 grid on their screen and communicated via a text box. Second, we did not use a trained confederate. We randomly assigned one of the players to the role of the instruction giver, and maintained the original wording by scripting a subset of their instructions. Finally, the hidden object was not placed in a bag, in which respect our design more closely resembles \cite{KeysarBarr___Brauner00_TakingPerspective}.

@@

In Table \ref{table:mainResults} we show the error rates on critical items, and compare to the data from
\cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
We find that 93\% of participants (all but one) attempted to move the hidden distractor at least once in the Experimental condition, out of four possible items, compared to only 7\% (only one) in the baseline condition. This is similar to the effect observed by the authors in the original study, which found 71\% and 0\%. Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
%Next, we examine whether there are differences in error rates across items. 
In Table \ref{table:ItemWise}, we break down the pattern of errors by item. 
%Because our randomization procedure undersampled the experimental condition for several items, a chi-squared test is inappropriate. Using Fisher's exact test, we cannot reject the null hypothesis of independence ($p = 0.13$). Still, 
We note that several items have much higher error rates than others -- for example, 75\% of participants in the experimental condition of item 6 made an error (the ``whiteboard eraser'' vs. the ``pencil eraser'') while only 17\% of participants in item 8 made an error (the ``computer mouse'' vs. the ``toy mouse''). Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target. We empirically substantiate this observation in our results for Expt.~2 below.

This item-wise variability suggests that the dependent variable highlighted in the original study (i.e. ``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item. Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants. 
%This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.

As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis. We define the decision window as the span of time between the point when the matcher received their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. Within the decision window, we computed the total amount of time spent hovering over the cell containing the target and divided by the total length of the decision window to get a measure of the relative time spent considering the target. We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. A paired-samples $t$-test found that people tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
We tested this prediction in Expt.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished.
By the reasoning above we expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.

% discuss other criticisms of Keysar paradigm?
While these results are compelling, the paradigm has been criticized from few different angles. 
\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well. 
We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
% Although relative fit was not empirically evaluated prior to our study,
Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.

Error rates are reported in Table \ref{table:mainResults}, alongside the results from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse} and our scripted replication in Expt.~1. Participants never moved the hidden object in the baseline condition, and total error rates for experimental trials are significantly lower than the rates found in Expt.~1, $\chi^2(1) = 5.35, p = 0.02$.

We find that patterns of errors in Expt.~2 diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$. Looking more closely at these patterns, we see that the only items where errors are consistently made are those where the more precise utterances used remain ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' from Expt.~1, but the two objects were still confusable at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total percentage of errors on experimental trials drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.

When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$. To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time. We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.

@@

After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials? We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.
% \begin{figure}[t!]
% \begin{center}
% \vspace{-.25cm}
% \includegraphics[scale=.55]{images/itemwisefitness.pdf}
% \vspace{-.75cm}
% \caption{Item-wise error rates compared with label fitness ratios, across both experiments. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
% \vspace{-.5cm}
% \label{fig:itemLevel}
% \end{center}
% \end{figure}

%\begin{itemize}
%\item people still try to move occluded objects, but they do it less, and that behavior is more confined to a few particularly difficult items.
%\item this happens because the scripted instructions we forced directors to use in Exp.~1 were unnatural and violated listener expectations of overinformativity
%\item the fact that the effect is attenuated in exp. 2 is actually evidence of some theory of mind use (insofar as the director choosing to be overinformative, and the matcher expecting overinformativity constitutes ToM
%\end{itemize}

@@

%\subsection{Discussion}
%
%In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it. Furthermore, the matcher tended to spend more time hovering their cursor over the occluded cell, and less time hovering over the target object, in ambiguous experimental trials than in baseline trials with no ambiguity. Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.
%
%Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First, 
%
%our analysis of item-level differences shows that the dependent variable highlighted in the original study (i.e. ``percentage of participants who moved the critical item at least once'') is somewhat problematic. It could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item. Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants. This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.

%Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers. Informally, we noted a tendency toward over-informativity. Instead of ``move the stuffed animal down'', they'll say ``move the stuffed pa'nda bear down.'' Or, instead of saying ``move the plane to the right,'' they'll say ``move the red airplane to the right,'' even though there is only one plane. This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations. Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment. If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells. We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished. If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.

%  \begin{figure}[b!]
% \begin{center}
% \vspace{-.5cm}
% \includegraphics[scale=.5]{images/mousetracking.pdf}
% \caption{Mean percentage of time spent hovering over target cell in the two conditions for Expt.~1 (left) and Expt.~2 (right). Error bars are boostrapped 95\% confidence intervals.}
% \vspace{-.5cm}
% \label{fig:hover}
% \end{center}
% \end{figure}

@@

It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room. We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.  On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication. Additionally, aspects of the interface such as the graphical representation of occluded cells may be less intuitive, or require more training, on the web than in the lab. % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.

>>>>

%While these failures are compelling evidence against a simple model where theory of mind reasoning is limited to literal perspective-taking, 
Past efforts to evaluate theory of mind in communication have focused on a relatively narrow behavioral operationalization of theory of mind as perspective-taking about a partner's visual access \citep{KeysarBarr___Brauner00_TakingPerspective, KeysarLinBarr03_LimitsOnTheoryOfMindUse, LinKeysarEpley10_ReflexivelyMindblind}. 
We argue in this paper that the mental models supporting theory of mind use in communication are much richer than previously explored. 
In particular, we consider a class of mental models representing not only a partner's visual access but also Gricean expectations of parsimony and cooperativity \citep{Grice75_LogicConversation}: that speakers will avoid saying things that are confusing or unnecessarily complicated given the current context. 
For example, suppose a friend wants to refer to a nearby Dalmatian. 
%There are numerous descriptions that would all be literally true, but different expressions are preferred in different contexts.
If there are no other dogs nearby, we may expect them to use a parsimonious basic-level term like ``dog.''
When there is potential for confusion, however, we expect them to consider our likely interpretation and introduce additional modifiers (``spotted dog'') or shift to more specific terms (``Dalmatian'') to disambiguate the one they are trying to refer to \citep{BrennanClark96_ConceptualPactsConversation, VanDeemter16_ComputationalModelsOfReferring, GrafEtAl16_BasicLevel}. 
When speakers violate these expectations, listeners draw inferences about the intended meaning in order to account for the violation: for example, XXX \cite{XXX, BrehenyKatsosWilliams06_ImplicaturesOnline, StillerGoodmanFrank15_AdHocImplicature}.

@@

% Discuss other criticisms of Keysar paradigm?
While these results are compelling, the paradigm has been criticized from few different angles. 
\cite{HellerGrodnerTanenhaus08_Perspective} have pointed out that in many cases, the hidden object was a better fit for the referring expression than the one in common ground (e.g. the hidden roll of tape vs. the cassette tape for ``the tape''), making the hidden object \emph{a priori} more likely to be the referent; it would generally be fairer to compare two objects that fit the referring expression equally well. 
We validate this argument by empirically measuring relative fit of the expressions to the target and distractor items.
Moreover, \cite{HannaTanenhausTrueswell03_CommonGroundPerspective, HannaTanenhaus04_PragmaticEffects} argued that the viewpoint asymmetry paradigm is somewhat unnatural: common ground is typically built incrementally over the course of an interaction rather than presented all at once, and it is rare for a shared display to differ in perceptual accessibility.

@@

%To deal with participants arriving at unpredictable times, we implemented a waiting room mechanism, where the first person to enter the game environment would see a screen saying ``Waiting for another player..." and the second person would trigger the game to start. 
Participants were randomly assigned to `director' and `matcher' roles and placed in a shared environment containing a chat box and a $4 \times 4$ grid of objects (see Fig. \ref{fig:interface}). Five of the grid cells were blocked from the director's view by a curtain, and six or seven objects were displayed in the grid at a given time. Trials were blocked into eight `items', each using a different set of objects. For each item, we gave the director a series of four instructions to move objects, which were displayed as a series of arrows pointing from some object to a neighboring unoccupied cell. They used a chat box to communicate the identity of this object and action to their partner, who proceeded to click and drag an object to complete the trial.

We used a between-subject design to test whether the scripted confederate labels were in fact what cooperative speakers would say in context. In the `scripted' condition, directors were free to communicate however they wished for half of the instructions but for the other half, including all the critical instructions, their messages were pre-scripted using the precise wording from \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. For example, when giving instructions on how to move the cassette tape, the director would be forced to use the ambiguous utterance ``Move the tape down one square." (That is, in these conditions, the scripted message would automatically appear in the director's chat box, and they would have to click `send' for the experiment to continue.) In the `unscripted' condition, directors were unrestricted and free to send whatever messages they deemed appropriate on all trials.

In each block, one target belonged to a `critical pair' of objects, such as a visible cassette tape and a hidden roll of tape that could both plausibly be referred to as `the tape'. We collected baseline performance for each condition by replacing the hidden alternative (e.g.~a roll of tape) with an object that did not fit the critical instruction (e.g.~a battery); we used the same unambiguous replacements as \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Each participant received half the items in the experimental condition and half in the baseline condition. The assignment of items to conditions was randomized across participants, and the order of conditions was randomized under the constraint that the same condition would not be used on more than two consecutive items. All object sets, object placements, and corresponding instruction sets were the same for all participants.

%On half of the instruction sets (the `critical' condition), the critical distractor object was placed in an occluded slot such that it was only visible to the matcher; on the other half (the `baseline' condition), all hidden distractors were unrelated fillers.

%Before entering the game environment, every participant independently passed a short quiz about the task's instructions, ensuring that they understood the interface. %Among other items on the quiz, we verified that both participants understood that items behind black cells were only visible to the matcher.

To collect clean mouse-tracking data, we asked the matcher to wait until the director sent a message; when the message was received, the matcher clicked a small circle in the center of the grid to continue the trial. We started recording from the matcher's mouse after this click.

@@

\subsubsection{Critical errors}

In Figure \ref{table:mainResults} we compare error rates across scripted and unscripted conditions. Participants rarely moved the hidden object on baseline trials in either condition, and total error rates for critical trials are significantly higher when speakers used scripted utterances, $\chi^2(1) = 5.35, p = 0.02$. Error rates in the scripted condition were similar to the effects observed by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. %, which found 71\% and 0\%. Our errors were larger across the board, perhaps due to the interface or the population, but the gap between the two conditions is roughly the same size.
%Next, we examine whether there are differences in error rates across items.

We also found substantial heterogeneity in errors rates across items (Table \ref{table:ItemWise}). Patterns of errors diverge significantly from a uniform distribution across items, $\chi^2(7) = 24.8, p < 0.001$.
For example, 75\% of participants made an error on item 6 upon hearing the scripted instruction ``the eraser'' while only 17\% of participants in item 8 made an error upon hearing ``the mouse.'' Informally, it seems as though the more difficult items are the ones where the utterance fits the distractor better than the target.  Looking more closely at these patterns, we see that the only items where errors are consistently made in the \emph{unscripted} condition are those where the more precise utterances used remain ambiguous. For example, in item 5, both the target and distractor are hair brushes: one is round and one is flat. Many participants produced the sub-class label ``hair brush,'' which was more precise than the scripted basic-level ``brush'' but the two objects were still confusable at the sub-class level. If we remove the two most difficult items (the ``hair brushes'' and the ``erasers''), the total error rate in the unscripted condition drops from 24\% to 10\% and the percentage of participants making at least one error drops from 61\% to 32\%.

%This item-wise variability suggests that the dependent variable highlighted in the original study (i.e. ``percentage of participants who moved the critical item at least once'') is somewhat problematic: it could look like 100\% of participants made errors even if they all made those errors on one particularly difficult item. Indeed, if we exclude the three `hard' items where over 60\% of participants in the experimental condition made errors, this dependent variable drops from 93\% to only 43\% of participants. 
%This suggests that the ``at least one error'' DV is not appropriate in settings with high variability across items and that we might want to be more careful about controlling for this variability, perhaps by first measuring salience and ambiguity without the occlusion aspect of the paradigm.

\subsubsection{Mouse-tracking}

As a proxy for the eye-tracking analyses reported by \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}, we conducted a mouse-tracking analysis. %We define the decision window as the span of time between  their instruction message and when they started moving an object. If it took them multiple attempts to move the correct object, we restricted our analysis to the first attempt. 
Within the decision window, we computed the total proportion of time spent moving toward the target relative to the distractor. %We had to exclude an additional 3 participants for this analysis, because the timestamps for director and matcher did not align and we could not establish the decision window properly. 
%A paired-samples $t$-test found that matchers in the scripted condition tended to spend less time hovering over the target cell on critical experimental trials than on baseline trials, $t(10) = -2.65, p = 0.02$ (see Fig. \ref{fig:hover}), indicating that the presence of a hidden distractor interfered with participants ability to directly choose the target.%\footnote{This analysis includes participants who actually made errors, since the data is too sparse to exclude them.} . %accounts for a significant proportion of variance $F(3, 52) = 10.7, p < 0.001$

%When we conducted a mouse-tracking analysis identical to the one reported for Expt.~1, we found no significant decrease in target hover time between experimental and baseline trials $t(27) = 0.89, p = 0.38$. To directly test for differences in hover time patterns across the two experiments, we used a mixed-effects model with a random intercept for game ID and an interaction between condition and experiment on target hover time. We found a marginally significant interaction, $b = 0.09, t = 1.89, p = 0.066$ (see Figure \ref{fig:hover}), providing some evidence that the presence of a hidden distractor no longer interfered target selection in Expt~2.

\subsubsection{Informativity}

By running this replication as a multi-player web experiment we have available an additional source of data beyond the original experiments: half of the instructions were unscripted, providing observations of natural production of referential descriptions for filler items. 
Informally, we noted a tendency toward additional, possibly unnecessary, precision in descriptions. Instead of ``move the stuffed animal down'', participants said ``move the stuffed panda bear down.'' Or, instead of saying ``move the plane to the right'' when there is only one plane, participants said ``move the red airplane to the right.'' 
Perhaps directors were taking the time to make more precise descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment increasing the chance of miscommunication from imprecise descriptions. 
If the matcher expected the director to be precise, then they would be justified in picking the first or best object that meets the description (rather than worrying excessively about occluded cells). 
That is, the scripted instructions used by the director for critical trials may have been \emph{uncooperative} for this situation, and thus led matchers astray. 
We expected to see more precise descriptions by unscripted directors and fewer errors by matchers in the critical trials.

@@

After separately establishing that matchers in Expt.~2 make fewer mistakes, that directors in Expt.~2 produce more informative utterances, and that informativity captures item-wise variability in error rates in both experiments, we tested the link between these effects in our aggregated data: does label informativity generally predict errors on critical trials? We used a mixed effects logistic regression model to estimate the effect of target and distractor fit on the probability of making a critical error in both experiments, including a random intercept for game ID. We found that participants are less likely to make errors when the target fit is higher, $b = -0.8, z = -4.1, p < 0.001$ and more likely to make errors when the distractor fit is higher, $b = 1.7, z = 3.7, p < 0.001$. Furthermore, a model including target fit and distractor fit in addition to item-level fixed effects is significantly better than a model including item alone, $\chi^2(2) = 36.2, p < 0.001$, implying that speaker informativity captures residual variance beyond the item-wise effects reported above.

% \begin{figure}[t!]
% \begin{center}
% \vspace{-.25cm}
% \includegraphics[scale=.55]{images/itemwisefitness.pdf}
% \vspace{-.75cm}
% \caption{Item-wise error rates compared with label fitness ratios, across both experiments. Error bars are bootstrapped 95\% confidence intervals for fitness and 95\% highest posterior density intervals for error rates.}
% \vspace{-.5cm}
% \label{fig:itemLevel}
% \end{center}
% \end{figure}

@@

\subsection{Discussion}

In both raw patterns of error and mouse-tracking metrics, we successfully replicated the findings of \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}. Participants frequently attempted to move the occluded alternative when it matched the target referential expression, even though the director couldn't possibly have been referring to it. Furthermore, the matcher tended to spend more time hovering their cursor over the occluded cell, and less time hovering over the target object, in ambiguous experimental trials than in baseline trials with no ambiguity. Both of these results support Keysar's claim that adults do not reliably deploy theory of mind reasoning when it is called for.

Still, in running this replication as a multi-player web experiment, we observed two particularly interesting features in the data. First,

Second, because half of the instructions were unscripted, we were able to observe natural production of referential descriptions for a number of fillers. Informally, we noted a tendency toward over-informativity. Instead of ``move the stuffed animal down'', they'll say ``move the stuffed pa'nda bear down.'' Or, instead of saying ``move the plane to the right,'' they'll say ``move the red airplane to the right,'' even though there is only one plane. This suggested that the confederate's scripted instructions were in some sense uncooperative and counter to the matcher's expectations. Perhaps directors were taking the time to make more costly over-informative descriptions because they believed it was contextually relevant: both parties know that there are hidden objects in the environment. If the matcher reasoned in this way and expected the director to be overinformative, then they would be justified in picking the first object that meets the description rather than worrying about occluded cells. We tested this prediction in Exp.~2, where we removed the scripted instructions and allowed speakers to refer to items however they wished. If speakers really are being overinformative, and listeners are correctly expecting this, we should see fewer errors after removing uncooperative scripts.

  \begin{figure}[t!]
 \begin{center}
 \vspace{-.5cm}
 \includegraphics[scale=.3]{figures/Exp2_design}
 \caption{Within-pair factorial design used in Exp.~2. In the common condition, we reproduce a simple pragmatic reference contrast: when more than one star is present in common ground, speakers are expected to be more informative. On hidden versions of these condition, the same set of objects is visible, but some additional squares are occluded. If pragmatic directors represent uncertainty over what the matcher sees behind these squares, we predict that they will produce more informative utterances.}
 \vspace{-.5cm}
 \label{fig:hover}
 \end{center}
 \end{figure}

\subsection{Experiment~2: Production in uncertain contexts}
\label{sec:Exp2}

We recruited K participants (K/2 pairs) from Amazon Mechanical Turk. All participants were native English speakers. %Three pairs were excluded for making 2 or more errors on non-critical items.

To more directly test speaker behavior across different contexts, we moved to a more systematic stimulus domain. These objects varied along three discrete features: \emph{shape}, \emph{texture}, and \emph{color}. Each feature took four discrete values, giving a space of 64 possible objects. We placed these objects in a smaller $3 \times 3$ grid, and used a pure \emph{reference game} setup where listeners had to click the intended referent as opposed to following a movement instruction.

To test how speakers flexibly modify their informativity in response to the uncertainty posed by

@@

It is worth noting several significant differences between our study and \cite{KeysarLinBarr03_LimitsOnTheoryOfMindUse}.
The primary difference between our study and the original, of course, is that it was run on the web with participants connected through a virtual environment, instead of face-to-face in a room. We believe we addressed the major concern about exploring theory of mind in web experiments -- that participants do not truly believe they are interacting with another human -- by allowing instantaneous, responsive, real-time interaction.  On the other hand, it is known that textual communication, as in our chat box, can differ from face-to-face verbal communication. Additionally, we omitted the procedural step of placing the hidden distractor in a paper bag, and some aspects of the interface such as the graphical representation of occluded cells may be less intuitive, or require more training, on the web than in the lab. % Directors watched objects move through the actual moment-by-moment mouse trajectories that matchers used to drag them, and matchers saw the actual timing of the messages sent by the director.
